### 2016-06-06 branches/Anthony-nonblocking/docs/intro.tex
non-blocking halo exchange version
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2897 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  intro.tex
%
%  Contains introduction and quick start for users.
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2014 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\subsection{Compilation}

Identify your local C compiler. The following example uses \texttt{gcc}
in serial, and \texttt{mpicc} in parallel. For MPI in particular,
local details may vary.

@@

First, compile the MPI stub library in the \texttt{mpi\_s}
directory. Do this by editing the Makefile, and checking the compiler
is appropriate. You should be able to build the library and run the
tests using, e.g.,:

@@

Now compile the main code in the \texttt{src} directory. Again
edit the Makefile to check the compiler. Compilation should
look like:

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  intro.tex
%
%  Contains introduction and quick start for users.
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2016 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

If you look in  the \texttt{trunk} directory, you should see
\begin{lstlisting}
$ ls trunk
config  LICENSE      mpi_s   src       tests  version.h
docs    Makefile.mk  README  targetDP  util
\end{lstlisting}

@@

\subsection{Configuration}

You will need to copy one of the existing configuration files from
the \texttt{config} directory to the top level directory. This file
sets all the relevant local details for compilation and so on. For
example:
\begin{lstlisting}
$ cp config/lunix-gcc-default.mk config.mk
\end{lstlisting}
Edit the new \texttt{config.mk} file to check, or adjust, the options
as required. You will need to identify your local C compiler. The
example uses \texttt{gcc}
in serial, and \texttt{mpicc} in parallel. For MPI in particular,
local details may vary. A quick way to compile the code is via
\begin{lstlisting}
$ cd tests
$ make compile-mpi-d3q19
\end{lstlisting}
This integrates a number of separate stages which are described
indivually in the following sections. The resulting executable
will be \texttt{./src/Ludwig.exe} at the top level.

\subsubsection{targetDP}

A \texttt{targetDP} layer is required by the code in all cases. It
allows compilation for both simple CPU systems and
for a number of accelerator systems using a single source code. This
will make the appropriate choices depending what has been specified
in the top-level \texttt{config.mk} file. From the top-level directory
\begin{lstlisting}
$ cd targetDP
$ make
\end{lstlisting}
should produce the appropriate \texttt{libtarget.a} library.

@@

First, compile the MPI stub library in the \texttt{mpi\_s}
directory. You should be able to build the library and run the
tests using, e.g.:

@@

Now compile the main code in the \texttt{src} directory.
Compilation should look like:

### 2016-06-06 branches/Anthony-nonblocking/targetDP/doc/introduction.tex
non-blocking halo exchange version
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2897 40876441-3496-0410-a613-a2fca35b2ae2

%\begin{comment}
%TO DO.
%\end{comment}
\section{Abstract}

The targetDP API provides an abstraction layer which allows
applications to \underline{t}arget \underline{D}ata
\underline{P}arallel hardware in a platform agnostic manner, by
abstracting the memory spaces and hierarchy of hardware
parallelism. Applications written using targetDP syntax are
performance portable: the same source code can be compiled for
different targets (where we currently support GPU accelerators and
modern multicore or manycore SIMD CPUs), without performance
overheads. The model is appropriate for abstracting the parallelism
contained within each compute node, and can be combined with, e.g. MPI
to allow use on systems containing multiple nodes. The targetDP API is
primarily aimed at the types of parallelism found in grid-based
applications, but may be applicable to a wider class of problems.
This document introduces the targetDP memory and execution models, and
specifies the syntax and functionality of the
interface. Implementation details are also provided, and a simple
example is given to demonstrate usage.

\section{Motivation}
It is becoming increasingly difficult for applications to exploit
modern computers, which continue to increase in complexity and
diversity with features including multicore/manycore CPUs, vector
floating point units, accelerators such as GPUs and non-uniform
distributed memory spaces.  From a scientist's perspective, it is not
only imperative to achieve performance, but also to retain
maintainability, sustainability and portability. The use of a
simplistic, well structured and clearly defined abstraction layer such
as targetDP can allow the programmer to express the scientific
problems in a way that will automatically achieve good performance
across the range of leading hardware solutions.

\section{Glossary}

>>>>

%\begin{comment}
%TO DO.
%\end{comment}

It is becoming increasingly difficult for applications to exploit
modern computers, which continue to increase in complexity and
diversity with features including multicore/manycore CPUs, vector
floating point units, accelerators such as GPUs and non-uniform
distributed memory spaces.  From a scientist's perspective, it is not
only imperative to achieve performance, but also to retain
maintainability, sustainability and portability. The use of a
simplistic, well structured and clearly defined abstraction layer such
as targetDP can allow the programmer to express the scientific
problems in a way that will automatically achieve good performance
across the range of leading hardware solutions.

The targetDP API (first introduced in \cite{hpcc}) provides an
abstraction layer which allows applications to \underline{t}arget
\underline{D}ata \underline{P}arallel hardware in a platform agnostic
manner, by abstracting the memory spaces and hierarchy of hardware
parallelism. Applications written using targetDP syntax are
performance portable: the same source code can be compiled for
different targets (where we currently support GPU accelerators and
modern multicore or manycore SIMD CPUs), without performance
overheads. The model is appropriate for abstracting the parallelism
contained within each compute node, and can be combined with, e.g. MPI
to allow use on systems containing multiple nodes. The targetDP API is
primarily aimed at the types of parallelism found in grid-based
applications, but may be applicable to a wider class of problems.

The targetDP memory and execution models are described in Chapters \ref{sec:exmodel} and \ref{sec:memmodel} respectively, and the document goers on to specify the memory management functionality in Chapter \ref{chap:memmanage} and data parallel execution functionality in Chapter  \ref{sec:dpexecution}. Implementation details are also provided for the existing C and CUDA versions of targetDP throughout these chapters. Finally, a simple
example is given to demonstrate usage in Chapter \ref{chapter:examples}.

\section*{Glossary}
